---
title: "Statistical Inference Project Part 1 - Simulation Exercise"
author: "Daniel DeWaters"
date: "11/7/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```


# Statistical Inference Course Project Part 1 - Simulation Exercise
## 1. Overview
In this project I will compare the exponential distribution with Central Limit Theorem. I will do 1,000 simulations investigating the distribution of averages of 40 exponentials with a lambda of 0.2.

## 2. Simulations
```{r simulations}
# Set the seed for randomly generated exponentials
set.seed(400)

# Parameters for simulation
lambda <- 0.2
n <- 40
sim <- 1:1000

# Simulate population
pop_sim <- data.frame(runs = sapply(sim, function(x){mean(rexp(n, lambda))}))

# Summarize simulation results
summary(pop_sim)
```

## 3. Sample Mean and Variance VS Theoretical Mean and Variance
```{r sampVStheo}
#Calculate the sample/theoretical mean, variance, and standard deviation
sample_mean <- mean(pop_sim$runs)
theoretical_mean <- 1/lambda

sample_variance <- var(pop_sim$runs)
theoretical_variance <- ((1/lambda)^2)/n

sample_sd = sd(pop_sim$runs)
theoretical_sd = (1/lambda)/sqrt(n)

# Display a list of the sample/theoretical mean and variance
rbind(sample_mean, theoretical_mean, sample_variance, theoretical_variance)

# Display 95% confidence interval
t.test(pop_sim$runs)$conf.int
```

The sample mean and variance are close to the theoretical mean and variance. The 95% confidence interval contains the theoretical mean. 

## 4. Distribution
```{r plots}

# Plot histogram of means
ggplot(data=pop_sim, aes(runs)) +
  geom_histogram(aes(y=..density.., fill=..density..), bins=30) +
  geom_density(aes(colour="a"), size=1) +
  stat_function(fun=dnorm, args=list(mean=sample_mean,sd=sample_sd),aes(color="b"), size=1) +
  geom_vline(aes(xintercept=sample_mean, color="a"), size=1) +
  geom_vline(aes(xintercept=theoretical_mean, color="b"), size=1) +
  scale_color_discrete(name = "Density", labels = c("Sample", "Theoretical")) +
  theme(legend.position=c(.92, .9)) +
  labs(title="Distibution of Means of 40 Exponentials", x="", y="Count")
```

The means of 40 exponentials over 1000 simulations are are centered and distributed closely to the normal distribution.