---
title: "Statistical Inference Course Project"
author: "Daniel DeWaters"
date: "11/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```


# Statistical Inference Course Project
## Part 1 - Simulation Exercise
### Overview
In this project I will compare the exponential distribution with Central Limit Theorem

### Simulations
```{r simulations}

```
###Sample Mean vs Theoretical Mean
###Sample Variance vs Theoretical Variance
###Distribution

## Part 2 - Tooth Growth
### Overview

ToothGrowth data set contains the result from an experiment studying the effect of vitamin C on tooth growth in 60 Guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

### Loading The Data
```{r loadData}
data("ToothGrowth")

d <- 
  ToothGrowth %>%
  mutate(dose = as.factor(dose))

d_means <- 
  ToothGrowth %>%
  mutate(dose=as.factor(dose)) %>%
  group_by(supp, dose) %>%
  summarise(mean_length = mean(len))
```

### Summarizing The Data
```{r toothPlots}
g <- ggplot(d_means, aes(dose, mean_length))
g + facet_grid(cols=vars(supp)) + geom_bar(stat="identity", aes(fill=dose)) +
    ggtitle("Mean Tooth Growth by Treatment and Dosage") +
    xlab("Dosage (in mg)") + ylab("Mean Tooth Length") + theme(legend.position="none")

g <- ggplot(d, aes(x=dose, y=len, group=dose))
g + geom_boxplot(aes(fill=dose)) + facet_grid(cols=vars(supp)) +
    ggtitle("Boxplots of Tooth Growth by Treatment and Dosage") +
    xlab("Dosage (in mg)") + ylab("Tooth Length") + theme(legend.position="none")
```

Exploring the data shows the following:
  1. Mean tooth length increases with increased dosage
  2. Mean tooth length is higher at 0.5mg and 1.0mg dosage
  3. Mean tooth length is similar at the 2.0 mg dosage

### Comparing Tooth Growth by Supplement and Dosage
```{r}
# Make vectors of tooth lengths resulting from each supplement/dosage combination
oj0.5 <- d[d$supp=="OJ" & d$dose==0.5, 1]
oj1.0 <- d[d$supp=="OJ" & d$dose==1, 1]
oj2.0 <- d[d$supp=="OJ" & d$dose==2, 1]
vc0.5 <- d[d$supp=="VC" & d$dose==0.5, 1]
vc1.0 <- d[d$supp=="VC" & d$dose==1, 1]
vc2.0 <- d[d$supp=="VC" & d$dose==2, 1]

# Perform t-tests on supplement pairs at each dosage
t.test(oj0.5, vc0.5, alternative="greater")
t.test(oj1.0, vc1.0, alternative="greater")
t.test(oj2.0, vc2.0, alternative="greater")


# Perform t test on OJ and VC at all dosage levels
t.test(len~supp, data=ToothGrowth, paired=FALSE, var.equal=FALSE)
```

### Conclusions